### 2022-05-18 main/chapter_2ljets.tex
expand atlas references and rjr section

% begin with physics content and results

Our main result is the content of Figure~\ref{fig:2ljets_summary}, which shows
the data in all main regions of this analysis, along with the regions' names,
and post-fit background expectations.
This figure illustrates the background-only part of a \heplikelihood
model which is extended with signals to test alternative models.

We use two signal models in the $\twoljets$ electroweak analysis.
One is an MSSM chargino-neutralino model (C1N2) which acts to produce
$\chargino{1}\textrm{--}\neutralino{2}$ pairs which decay via $W$ and $Z$
channels respectively to jets and a lepton pair.
The other is a GMSB model which decays via Higgs and $Z$ bosons in variable
branching fractions.
Diagrams illustrating the C1N2 and GMSB signal models we consider are shown in
Figure~\ref{fig:2ljets_signal_diagrams}.
Some example effects of these signals in signal regions are displayed in
Figure~\ref{fig:2ljets_signal_examples}.

At each parameter point of a signal model, it is added to the background model
and tested against data in the $\mathrm{CLs}$ prescription.
Contours showing which regions where those tests label the signal model
parameters as excluded excluded are shown for
C1N2 in Figure~\ref{fig:2ljets_contours_c1n2} and for
GMSB in~\ref{fig:2ljets_contours_gmsb}.

@@

\begin{figure}[tp]
\centering
\includegraphics[width=0.99\textwidth]{figures/2ljets_contours_c1n2.pdf}
\caption{%
Contours for the C1N2 model in the $\twoljets$ electroweak analysis.
Space below the solid red line is labelled as excluded and its dotted
neighbours show the result if all signal cross-sections are varied up and down
by theoretical error bars.
The yellow band shows the $\pm1$-sigma region of exclusion contours
from asymptotic approximations to the prior distribution of the test statistic.
Grey areas are observed limits from the two-lepton parts
of~\cite{SUSY-2016-24} and~\cite{SUSY-2013-11}.
Exclusion is defined by the $95\%$ $\mathrm{CLs}$ prescription
in asymptotic approximations.
All contours are interpolated from a sparse grid.
}
\label{fig:2ljets_contours_c1n2}
\end{figure}

@@

\paragraph{Contributions}
My contributions the the analysis are listed here.
\begin{itemize}
\item Design, implementation and execution of the electroweak part of the analysis.
\item Service as `analysis contact' from January to October 2020.
\item Production of main data inputs for all three analyses
and the RJR $3\ell$ search \TODO{cite}.
Production of all systematic variations for the background samples;
other group members assisted with the central background sample.
Production of all electroweak signal samples and their systematic variations.
\item Integration with ATLAS combinations and pMSSM scan efforts by
performing their validation studies and serializing the analysis results.
\item Preparation of electroweak results for publication in the paper \TODO{cite} and the
public HEPData record~\cite{maguire2017hepdata}.
\end{itemize}
Except where specified, I produced all figures presented in this thesis.
Many plotting scripts, however, are modified from versions inherited from
many \atlas\ members past and present to whom I am grateful.

@@

\clearpage
\FloatBarrier
\section{Context}
% previous results on 2(/3)L from ATLAS
% other constraints on these models from ATLAS/CMS
% previous region design, motivation for this work
Previous searches studied this two-lepton + jets + $\ptmiss$ selection
in different or smaller data-sets in \atlas.
This project serves to update those results with the full LHC Run 2 data-set,
which contains $139~\mathrm{fb}^{-1}$ of usable proton-proton collisions at
$13~\eV[T]$ collision energy.

Increased quantity of data allows more options for this search.
Together with lessons learned from the previous searches, this is not only
an update but a chance for improvement.

This paper focusses solely on $\twoljets$ data and analyses them in the three
distinct RJR, electroweak, and strong parts.
Each part has precedent from previous searches which were conducted and
published separately with other work.
A decision was made for full Run 2 to group the searches by targetted physics,
rather than by method or targetted physical models.

The remainder of this section describes the specific history and motivations
for each of the RJR, electroweak and strong sub-projects in this search.

\subsection{RJR}
Recursive Jigsaw Reconstruction (RJR), is a method for
interpreting particle data as arising from user-specified decay trees.
The $\twoljets$ RJR analysis uses two C1N2 decay trees which are displayed in
Figure~\ref{fig:2ljets_rjr_decay_trees}.

\begin{figure}[tp]
\centering
\includegraphics[width=0.99\textwidth]{figures/2ljets_rjr_trees.pdf}
\caption{%
Decay trees for the $\twoljets$ RJR analysis.
Both diagrams represent processes also targetted in the electroweak analysis.
(left) A fully-resolved decay tree with its particles labelled.
(right) Only the $Z\rightarrow\ell\ell$ (\underline{v}ector boson)
is resolved, but the supersymmetric system recoils off
hard QCD jets labelled ``ISR'', boosting the \underline{i}nvisible particles,
to generate visibly large $\met$ in models with small mass-splittings.
The same approach is used for SR-OffShell selections in the electroweak
analysis.
}
\label{fig:2ljets_contours_gmsb}
\end{figure}

\subsection{Electroweak}

\subsection{Strong}

>>>>

% begin with physics content and results
My main result in this thesis is displayed in Figure~\ref{fig:2ljets_summary}.

It shows unblinded regions from the electroweak part of
Reference~\cite{atlas2022searches}, which reports searches
``for new phenomena in events with two leptons, jets, and missing
transverse momentum''.
For brevity, I will name that search $\twoljets$.

This section begins with a splash summary of interpretations of the results in
Figure~\ref{fig:2ljets_summary},
before cutting through the gory details later on.

In a large number orthogonal signal regions, where non-standard signals would
encourage excess data above backgrounds,
no significant excess is observed above \emph{post-fit} background expectations.
Indeed there are some significant deficits --- data below the background noise.

Background parts of a \heplikelihood model are illustrated behind data.
To test non-standard (supersymmetric) models against observation, signal
contributions are added on top of those backgrounds.

We use two signal models in the $\twoljets$ electroweak analysis.
One is an MSSM chargino-neutralino model (C1N2) which acts to produce
$\chargino{1}\textrm{--}\neutralino{2}$ pairs, which decay via
$W\rightarrow q\bar q$ and
$Z\rightarrow \ell^\pm \ell^\mp$ respectively..
The other is a GMSB model which, after a variety of initial states produces
$\neutralino{1}\textrm{--}\neutralino{1}$ pairs.
These neutralinos decay,
via Higgs and $Z$ bosons in variable branching fractions,
to final states again containing $\ell^\pm \ell^\mp$ and jet-seeding partons.
Diagrams illustrating these C1N2 and GMSB signal models are shown in
Figure~\ref{fig:2ljets_signal_diagrams}.

Effects of some of these these signals in example signal regions of the
$\twoljets$ electroweak analysis are displayed in
Figure~\ref{fig:2ljets_signal_examples}.

At each parameter point of a signal model, it is added to the background model
and tested against data in the $\mathrm{CLs}$ prescription.
Contours showing which parameters of those signal models are labelled as
excluded are displayed for
C1N2 in Figure~\ref{fig:2ljets_contours_c1n2} and for
GMSB in~\ref{fig:2ljets_contours_gmsb}.

@@

\begin{figure}[tp]
\centering
\includegraphics[width=0.99\textwidth]{figures/2ljets_contours_c1n2.pdf}
\caption{%
Contours for the C1N2 model in the $\twoljets$ electroweak analysis.
Space below the solid red line is labelled as excluded and its dotted
neighbours show the result if all signal cross-sections are varied up and down
by theoretical error bars.
The yellow band shows the $\pm1$-sigma region of exclusion contours
from asymptotic approximations to the prior distribution of the test statistic.
Grey areas are observed limits from the two-lepton parts
of~\cite{atlas_23l_SUSY_2016_24} and~\cite{atlas_2l_SUSY_2013_11}.
Exclusion is defined by the $95\%$ $\mathrm{CLs}$ prescription
in asymptotic approximations.
All contours are interpolated from a sparse grid.
}
\label{fig:2ljets_contours_c1n2}
\end{figure}

@@

\FloatBarrier
\subsection{Contributions}
My contributions the the analysis are listed here.
\begin{itemize}
\item Design, implementation and execution of the electroweak part of the analysis.
\item Service as `analysis contact' from January to October 2020.
\item Production of main data inputs for all three analyses
and the RJR $3\ell$ search \TODO{cite}.
Production of all systematic variations for the background samples;
other group members assisted with the central background sample.
Production of all electroweak signal samples and their systematic variations.
\item Integration with ATLAS combinations and pMSSM scan efforts by
performing their validation studies and serializing the analysis results.
\item Preparation of electroweak results for publication in the paper \TODO{cite} and the
public HEPData record~\cite{maguire2017hepdata}.
\end{itemize}
Except where specified, I produced all figures presented in this thesis.
Many plotting scripts, however, are modified from versions inherited from
many \atlas\ members past and present to whom I am grateful.

@@

\clearpage
\FloatBarrier
\section{Origins}
% previous results on 2(/3)L from ATLAS
% other constraints on these models from ATLAS/CMS
% previous region design, motivation for this work
Previous analyses of \atlas\ data have explored similar selections of
two leptons, jets, and missing transverse momentum.
Quite a few analyses, in fact.
This project exists to supplement those results with information from the full
LHC Run 2 data-set which was completed in 2018 and contains
$139~\mathrm{fb}^{-1}$ of usable proton-proton collisions at a collision energy
of $13~\eV[T]$.
Leverage of the increased data quantity is one key aim.
Analysis quality also might be improved if our designs can take lessons from
the experiences of previous work.

% electroweak
% Run 1 data~\cite{atlas_2l_SUSY_2013_11},
% partial Run 2 data~\cite{atlas_23l_SUSY_2016_24},
% atlas_rjr_23l_SUSY_2017_03,
% atlas_rjr_mimic_SUSY_2018_06

The $\twoljets$ analysis is a trinity,
of which all three parts expand upon \atlas\ results which previously appeared
in separate publications.
These three parts are named `RJR', `electroweak', and `strong', each of which
is described in its historical context in this section.

Recursive Jigsaw Reconstruction (RJR) is an algorithm for constructing useful
event variables.
The $\twoljets$ RJR analysis, detailed in Section~\ref{sec:2ljets_origins_rjr},
uses RJR variables to define its regions in a signal-independent search.

My primary contributions are to the $\twoljets$ electroweak analysis,
detailed in Section~\ref{sec:2ljets_origins_electroweak},
which uses conventional (non-RJR) event variables to define a collection of
orthogonal regions in a search for effects from the electroweak sector of
supersymmetric models.

Unsurprisingly, the strong part of the $\twoljets$ analysis,
detailed in Section~\ref{sec:2ljets_origins_strong},
searches for effects from the strong sector.
It also uses conventional event variables and with an emphasis on the mass
spectrum of lepton pairs.

\subsection{RJR}
\label{sec:2ljets_origins_rjr}

\begin{figure}[tp]
\centering
\includegraphics[width=0.9\textwidth]{figures/2ljets_rjr_trees.pdf}
\caption{%
Decay trees for the $\twoljets$ RJR analysis.
Both diagrams represent processes also targetted in the electroweak analysis.
(left) A fully-resolved decay tree with its particles labelled.
(right) Only the $Z\rightarrow\ell\ell$ (\underline{v}ector boson)
is resolved, but the supersymmetric system recoils off
hard QCD jets labelled ``ISR'', boosting the \underline{i}nvisible particles,
to generate visibly large $\met$ in models with small mass-splittings.
The same approach is used for SR-OffShell selections in the electroweak
analysis.
\emph{This figure was made by the RJR team.}
}
\label{fig:2ljets_rjr_decay_trees}
\end{figure}

The Recursive Jigsaw Reconstruction (RJR) method interprets reconstructed
particle data by reconciling them with user-specified decay trees, and works
by making decisions of how to assign four-momenta to different `systems',
which are nodes of the chosen the decay graph.
Event variables are then evaluated in those systems' rest
frames~\cite{jackson2017sparticles, jackson2017rjr}.

In this way, RJR variables attain intelligence in how parent particles are
reconstructed from their visible decay products and approximate independence
from uninteresting boosts, particularly those from the emission of QCD jets
outside of the supersymmetric decay tree.

Over the last several years, RJR event variables have been demonstrated as
effective tools for interpreting supersymmetric particle physics
data~\cite{santoni2018probing},
have been used in numerous \atlas\ analyses~\cite{%
atlas_rjr_SUSY_2016_07,
atlas_rjr_SUSY_2016_15,
atlas_rjr_SUSY_2016_16,
atlas_rjr_23l_SUSY_2017_03
}, and have influenced the designs of other
analyses~\cite{atlas_rjr_mimic_SUSY_2018_06}.

Decay trees used in the $\twoljets$ RJR analysis,
which displayed in Figure~\ref{fig:2ljets_rjr_decay_tres},
target C1N2-like scenarios similar to those also considered in the $\twoljets$
electroweak search.

\subsection{Electroweak}
\label{sec:2ljets_origins_electroweak}

\subsection{Strong}
\label{sec:2ljets_origins_strong}
